---
title: "Central Limit Therorem on an Exponential Distribution"
author: "Justin Pizzino"
date: "July 5, 2019"
output: 
  pdf_document:
    latex_engine: pdflatex

---
# Summary:
### Below I will prove the cental limit therorem and show how sampling a population can create a normal distribution for data that may not be normal. 

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
  library(R.utils)
  library(data.table)
  library(dplyr)
  library(lubridate)
  library(ggplot2)
```

# Create the simulation & plot the results
```{r simulation }
  rate<-0.2
  n<-40
  intervals<-1000
  data = NULL
  for (i in 1 : intervals) data = c(data, mean(rexp(n,rate)))
```
### The sample mean is: `r mean(data)` (also shown in blue below), which is very close to the therdoical mean of 1/lamda or 1/.2 = 5.

```{r plot, echo=FALSE}

data<-data.table(data)
g<-ggplot(data, aes(x=data)) 
g<-g + geom_histogram(color="black",fill="white", bins=30)
g<-g + geom_vline(aes(xintercept=mean(data)),
                  color="blue",linetype="dashed",size=1)
g<-g + theme(axis.text.x=element_text(size=12, color="#0000ff")) 
g<-g+ scale_x_continuous(breaks=c(3:8))
g<-g+labs(title="Sample Distribution with Mean", x="Value",y="Count")
print(g)
```

### Now to compare sample variance to the theoretical variance

##### Popultion or theoretical variance is mean^2/n so in this case 0.04/40: `r (1/(.2)^2)/40`

##### The sample vairance using r function var() is: `r var(data)`
