---
title: "Central Limit Theorem on an Exponential Distribution"
author: "Justin Pizzino"
date: "July 5, 2019"
output: 
  pdf_document:
    latex_engine: pdflatex

---
# 1. Summary:
Below I will prove the central limit theorem and show how sampling a population can create a normal distribution for data that may not be normal. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
  library(R.utils)
  library(data.table)
  library(dplyr)
  library(lubridate)
  library(ggplot2)
```

# 2. Create the simulation & plot the results
I used a for loop to run the rexp function to calculate 40 sample size simulation.  I ran this 1000 times and added the means to a data table.
```{r simulation }
  set.seed(8534)
  rate<-0.2
  n<-40
  intervals<-1000
  sim<-NULL
  for (i in 1 : intervals) sim = c(sim, mean(rexp(n,rate)))
  
  
  sim<-data.table(sim)
  theoMean<-(1/rate)
  theoVar<-theoMean^2/n
  sampMean<-mean(sim$sim)
```

Below is a chart showing the mean of the 1000 simulations.  The mean is the sample is near 5 which is close to our theoretical mean.

```{r plotCount, echo=FALSE, fig.width=11, fig.height=5}
g<-ggplot(sim, aes(x=sim)) 
g<-g + geom_histogram(color="black",fill="white", bins=30)
g<-g + geom_vline(aes(xintercept=mean(sim)),
                  color="blue",linetype="dashed",size=1)
g<-g + theme(axis.text.x=element_text(size=12, color="#0000ff")) 
g<-g+ scale_x_continuous(breaks=c(3:8))
g<-g+labs(title="Sample Distribution with Mean", x="Mean",y="Count")
g<-g+theme(axis.text = element_text(size=16), axis.title = element_text(size=16))
g
```

# 3. Now to compare sample variance & mean to the theoretical variance & mean, as well as review confidence levels

**Confidence interval:** `r t.test(sim$sim)[4]`

**The sample mean is:** `r mean(sim$sim)` (also shown in blue below), which is very close to the theoretical mean of 1/lamda or 1/.2 = 5.

**Theoretical variance** is mean^2/n so in this case 0.04/40: `r (1/(.2)^2)/40`

**The sample variance** using r function var() is: `r var(sim)`
These values are very close, and the graph below shows how close the two distributions are (red and green curves)

# 4. Distributions
Let's look at the distribution to show it is normal.  As can be seen below the green curve appears symmetric bell shaped, with ~68% of the values with in 1 standard deviation.


```{r plotMeans, echo=FALSE, fig.width=11, fig.height=5}
g2<-ggplot(sim, aes(x=sim))
g2<-g2 + geom_histogram(aes(y=..density..),color="black",fill="white", bins=30)
g2<- g2 + stat_function(fun=dnorm, args = list(mean=theoMean, sd=sqrt(theoVar)), color="red")
g2<-g2 + labs(title="Density of 40 Exponential with 1k Simulations", x="Mean",y="Density")
g2<-g2+geom_density(color="green")
g2<-g2 + geom_vline(aes(xintercept=mean(sim)),
                  color="green",linetype="dashed",size=1)
g2<-g2 + geom_vline(aes(xintercept=theoMean),
                  color="red",linetype="dashed",size=1)
g2<-g2+theme(legend.position="right")
g2<-g2+theme(axis.text = element_text(size=16), axis.title = element_text(size=16))
g2
```

The red shows sample mean and distribution, green shows theoretical.  As Central Limit Theorem states that sample means of simple random elements from a large (infinite) population will take on a bell shape.  It is acceptable to have 30 or more in a sample and we have 40 samples by 1000 simulations.  As you can see here our distribution is bell shaped and is close to the theoretical.

# 5. Does the distribution means of 40 exponential behave as predicted by he Central Limit Theorem?
**CLT states** - the distribution of averages of iid variables (properly normalized) becomes that of a standard normal as the sample size increases.

**Yes**, based on the simulation run, the comparison of theoretical to sample, and charts displayed, the samples do behave as predicted by CLT.  The sample means as charted appear normal.  **Also the sample and theoretical means and variances are close to each other as predicted. **

# 6. Appendix

Plot 1 code
```{r plotCount2, echo=TRUE, eval = FALSE}
g<-ggplot(sim, aes(x=sim)) 
g<-g + geom_histogram(color="black",fill="white", bins=30)
g<-g + geom_vline(aes(xintercept=mean(sim)),
                  color="blue",linetype="dashed",size=1)
g<-g + theme(axis.text.x=element_text(size=12, color="#0000ff")) 
g<-g+ scale_x_continuous(breaks=c(3:8))
g<-g+labs(title="Sample Distribution with Mean", x="Mean",y="Count")
print(g)
```


Plot 2 code
```{r plotMeans2, echo=TRUE, eval = FALSE}
g2<-ggplot(sim, aes(x=sim))
g2<-g2 + geom_histogram(aes(y=..density..),color="black",fill="white", bins=30)
g2<- g2 + stat_function(fun=dnorm, args = list(mean=theoMean, sd=sqrt(theoVar)), 
                        color="red")
g2<-g2 + labs(title="Density of 40 Exponential with 1k Simulations", x="Mean",
              y="Density")
g2<-g2+geom_density(color="green")
g2<-g2 + geom_vline(aes(xintercept=mean(sim)),
                  color="green",linetype="dashed",size=1)
g2<-g2 + geom_vline(aes(xintercept=theoMean),
                  color="red",linetype="dashed",size=1)
g2<-g2+theme(legend.position="right")
print(g2)
```