---
title: "Tooth Growth Analysis"
author: "Justin Pizzino"
date: "7/7/2019"
output: 
  pdf_document:
    latex_engine: pdflatex

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
  library(R.utils)
  library(data.table)
  library(dplyr)
  library(lubridate)
  library(ggplot2)
  library(datasets)
  options(scipen=999)
```
# 1. Summary
Using the r Tooth Growth data I will make a judgement as to which delivery method and dose resulted in the most growth.  This was done through plotting the results, and reviewing the t.test for the different methods and dosages.

# 2. Load the data and provide a summary
```{r load, echo=TRUE}
data(ToothGrowth)
summary(ToothGrowth)
```
## Let's review the two delivery methods, VC and OJ
Mean of Tooth Growth by VC: `r mean(subset(ToothGrowth,supp=="VC")[,"len"])` and by OJ `r mean(subset(ToothGrowth,supp=="OJ")[,"len"])`

By plotting the data we can see that tooth growth increases consistently using VC, but tapers off a bit after 1.0 dose with OJ.

```{r plot, echo=FALSE, fig.width=11, fig.height=5}
g<-ggplot(ToothGrowth, aes(dose, len))
g<-g+facet_grid(.~ supp)+geom_point(color="red")
g<-g+stat_summary(fun.y=mean, geom="line")
g<-g+labs(title="Comparison of Tooth Growth by Delivery Method", x="Dose", y="Length")
g<-g+theme(axis.text = element_text(size=16), axis.title = element_text(size=16))
g
```

## Let's review the different dose types, 0.5, 1.0, 1.5, and 2.0
```{r dose, echo=FALSE, fig.width=11, fig.height=5}
tgSum<-summarise(group_by(ToothGrowth,dose), Avg_Len=mean(len))
g2<-ggplot(tgSum, aes(x=dose, y=Avg_Len))+geom_line()
g2<-g2+labs(title="Average Length by Dose", x="Dose", y="Average Length")
g2<-g2+theme(axis.text = element_text(size=16), axis.title = element_text(size=16))
g2

```
We can see here that a higher dosage seem to improve length, but starts to taper off at 1.0.

# 3. Use confidence intervals and/or hypothesis test to compare tooth growth by supp and dose
Using a 95% confidence level let's run some tests on the data.

Below we see that the mean of OJ is higher but the p-value is 0.06 which is higher than 0.05 resulting in not enough evidence to reject the null hypothesis. 
```{r echo=TRUE}
t.test(subset(ToothGrowth,supp=="VC")[,"len"],subset(ToothGrowth,supp=="OJ")[,"len"], paried=FALSE, var.equal=FALSE)
```

Now let's compare the different doses, starting with 0.5 to 1.0.  Here you can see the mean is higher for 1.0 dose, and p-value is much lower, and the confidence interval doesn't include 0.  Thus we can reject the null hypothesis
```{r echo=TRUE}
t.test(subset(ToothGrowth,dose==0.5)[,"len"],subset(ToothGrowth,dose==1.0)[,"len"], paried=FALSE, var.equal=FALSE)
```
Compare 1.0 to 2.0
```{r echo=TRUE}
t.test(subset(ToothGrowth,dose==1.0)[,"len"],subset(ToothGrowth,dose==2.0)[,"len"], paried=FALSE, var.equal=FALSE)
```
Here we can see the mean for 2.0 is higher then 1.0, but the change is less so then from 0.5 to 1.0.  Also the p value is lower, but still, we cannot reject the null hypothesis as the confidence interval does not contain 0, and we have a very low p-value.

# 4. State your conclusions and the assumptions needed for your conclusions
**Do delivery method and/or dosage affect tooth growth in guinea pigs?**
Using the data provided and assuming there are no other factors in the different methods of supplement deliver and dosage I conclude that:

1) OJ delivery method may provide better results at 0.5 and 1.0 dosage, but there is not enough data to support that holistically, further more at a dosage of 2.0 the difference is no longer there.

2) Tooth growth increases at a dosage of 1.0 over 0.5.  Tooth growth continues to increase at a dosage of 2.0, but tappers off.

# 5. Appendix
Code for plot 1
```{r plot_code, echo=TRUE, eval=FALSE}
g<-ggplot(ToothGrowth, aes(dose, len))
g<-g+facet_grid(.~ supp)+geom_point(color="pink")
g<-g+stat_summary(fun.y=mean, geom="line")
g<-g+labs(title="Comparison of Tooth Growth by Delivery Method", x="Dose", y="Length")
g
```
Code for plot 2
```{r dose_code, echo=TRUE, eval=FALSE}
tgSum<-summarise(group_by(ToothGrowth,dose), Avg_Len=mean(len))
g2<-ggplot(tgSum, aes(x=dose, y=Avg_Len))+geom_line()
g2<-g2+labs(title="Average Length by Dose", x="Dose", y="Average Length")
g2

```